%% ============================================================
%% shared/services/storage/s3.tex
%% Amazon S3 — SAA + SAP
%% ============================================================

\servicesection{Amazon S3}{\bothtag}{Object storage --- 11 nines durability, unlimited capacity}

\begin{keyfacts}
\begin{itemize}
  \item Objects stored in \textbf{buckets}; bucket names are globally unique.
  \item Object size: 0\,bytes to 5\,TB; max single PUT is 5\,GB.
        Use \textbf{Multipart Upload} for objects $>$100\,MB.
  \item \textbf{Durability}: 99.999999999\% (11 nines) --- replicated across $\geq$3 AZs.
  \item \textbf{Availability}: varies by storage class (Standard = 99.99\%).
  \item S3 is a \textbf{regional} service; bucket name namespace is global.
  \item \textbf{Strong consistency} (since Dec 2020): all operations are strongly consistent.
  \item Throughput: \textbf{3\,500 PUT/COPY/DELETE} + \textbf{5\,500 GET/HEAD} per prefix per second.
\end{itemize}
\end{keyfacts}

\subsection{Storage Classes}

\begin{comparebox}[title={S3 storage class quick reference}]
\begin{tabular}{@{}p{3.2cm}p{2cm}p{1.5cm}p{4cm}@{}}
\toprule
\textbf{Class} & \textbf{Retrieval} & \textbf{Min days} & \textbf{Use case} \\
\midrule
Standard               & Milliseconds & ---  & Frequently accessed data \\
Intelligent-Tiering    & Milliseconds & ---  & Unknown / changing access patterns \\
Standard-IA            & Milliseconds & 30   & Infrequent access, fast retrieval \\
One Zone-IA            & Milliseconds & 30   & Reproducible infrequent data (single AZ) \\
Glacier Instant        & Milliseconds & 90   & Archives needing instant access \\
Glacier Flexible       & 1 min -- 12 h & 90  & Backup / archive, hours OK \\
Glacier Deep Archive   & 12 h         & 180  & Long-term compliance, cheapest \\
\bottomrule
\end{tabular}
\end{comparebox}

\begin{examtip}
Retrieval fees apply to IA and Glacier classes. Glacier Deep Archive = cheapest storage but
12-hour retrieval --- wrong for ``retrievable within hours'' scenarios (use Glacier Flexible).
One Zone-IA loses data if that AZ fails --- never for unique or non-reproducible data.
\end{examtip}

\subsection{Security}

\begin{itemize}
  \item \textbf{Block Public Access}: account-level or bucket-level setting;
        overrides any bucket policy or ACL. Enable by default unless hosting public static sites.
  \item \textbf{Bucket Policies}: resource-based JSON IAM policies; can grant cross-account access.
  \item \textbf{ACLs}: legacy; AWS recommends disabling ACLs and using bucket policies instead.
  \item \textbf{Pre-signed URLs}: time-limited URL grants temporary access to a private object
        without changing bucket permissions. Generated by a principal with \texttt{s3:GetObject}.
  \item \textbf{VPC Gateway Endpoint}: route S3 traffic within VPC without NAT / internet gateway.
        Free; add to route tables. (Interface Endpoint = PrivateLink, costs per GB.)
\end{itemize}

\subsubsection{Server-Side Encryption (SSE)}

\begin{comparebox}[title={S3 encryption options}]
\begin{tabular}{@{}p{2.5cm}p{4cm}p{5cm}@{}}
\toprule
\textbf{Type} & \textbf{Key managed by} & \textbf{Notes} \\
\midrule
SSE-S3  & AWS (S3-managed AES-256) & Default; no extra cost \\
SSE-KMS & AWS KMS (CMK or AWS-managed) & Audit trail via CloudTrail; extra KMS API cost \\
SSE-C   & Customer (you supply key per request) & AWS does not store key; HTTPS required \\
Client-side & Customer (encrypt before upload) & Full key control; more complex \\
\bottomrule
\end{tabular}
\end{comparebox}

\begin{examtip}
\textbf{Default encryption}: enable SSE-S3 or SSE-KMS on the bucket so all new objects
are encrypted automatically. To \emph{require} encryption, add a bucket policy that
\texttt{Deny}s PUT requests without the \texttt{x-amz-server-side-encryption} header.
\end{examtip}

\subsection{Versioning \& Replication}

\textbf{Versioning}
\begin{itemize}
  \item Once enabled, can only be \emph{suspended}, never disabled.
  \item Stores all versions of an object (including deletes as delete markers).
  \item \textbf{MFA Delete}: requires MFA to permanently delete versions or change versioning state.
\end{itemize}

\textbf{Replication}
\begin{itemize}
  \item \textbf{CRR} (Cross-Region Replication): DR, compliance, latency.
  \item \textbf{SRR} (Same-Region Replication): log aggregation, test/prod sync.
  \item Both require \emph{versioning enabled} on source and destination buckets.
  \item Replication is \emph{not retroactive}: existing objects are not replicated.
        Use \textbf{S3 Batch Operations} to replicate existing objects.
  \item Replication does not copy delete markers by default (configurable).
\end{itemize}

\subsection{S3 Lifecycle Policies}

Automate moving or deleting objects based on age:
\begin{itemize}
  \item \textbf{Transition actions}: move to cheaper storage class after N days
        (e.g.\ Standard $\rightarrow$ Standard-IA after 30 days $\rightarrow$ Glacier after 90 days).
  \item \textbf{Expiration actions}: delete objects (or noncurrent versions) after N days.
  \item Minimum 30-day rule: objects must stay in Standard-IA / One Zone-IA $\geq$30 days before transitioning.
\end{itemize}

\subsection{Performance \& Transfer}

\begin{itemize}
  \item \textbf{Multipart Upload}: parallel upload of parts; required for $>$5\,GB, recommended $>$100\,MB.
        Increases throughput, recovers from network failures.
  \item \textbf{S3 Transfer Acceleration}: routes uploads/downloads through CloudFront edge
        locations over AWS backbone. Good for global uploads to a single bucket.
  \item \textbf{S3 Select / Glacier Select}: run SQL queries against CSV/JSON/Parquet objects
        in S3; retrieve subset of data (up to 400\% faster, 80\% cheaper than full retrieval).
  \item \textbf{Byte-range fetches}: download only a specific byte range; enables parallel
        multi-thread downloads and partial object retrieval.
  \item Prefix performance: spread objects across prefixes to maximise throughput
        (5\,500 GET/prefix/s $\times$ many prefixes).
\end{itemize}

\begin{saadepth}
\textbf{SAA: S3 features for application architectures:}
\begin{itemize}
  \item \textbf{Static website hosting}: host HTML/CSS/JS directly from S3 (enable in bucket
        settings); pair with CloudFront for HTTPS and global CDN.
  \item \textbf{Event Notifications}: trigger SQS, SNS, Lambda, or EventBridge on
        \texttt{s3:ObjectCreated}, \texttt{s3:ObjectRemoved}, etc.
  \item \textbf{Object Lock (WORM)}: write-once-read-many; Governance mode (admins can override)
        vs Compliance mode (nobody can delete during retention, including root).
  \item \textbf{Requester Pays}: download costs billed to the requester, not bucket owner.
        Useful for data sets shared with external parties.
  \item \textbf{CORS}: configure cross-origin resource sharing so browser apps on domain A
        can fetch S3 objects hosted under domain B.
\end{itemize}
\end{saadepth}

\begin{sapdepth}
\textbf{SAP: S3 at enterprise scale:}
\begin{itemize}
  \item \textbf{S3 Access Points}: create named access points with their own policies for
        different applications or teams --- simplifies bucket policy management at scale.
  \item \textbf{Multi-Region Access Points}: single global endpoint routes requests to the
        lowest-latency bucket; failover across regions; enables active-active S3.
  \item \textbf{Object Lambda}: attach a Lambda function to a \texttt{GetObject} call;
        transform data on the fly (redact PII, convert format) without storing multiple copies.
  \item \textbf{S3 as Data Lake}: backbone for Athena (serverless SQL), AWS Glue (ETL),
        Lake Formation (governance, fine-grained access). Store raw + processed tiers.
  \item \textbf{S3 Batch Operations}: run Lambda or S3-native operations (copy, tag, restore,
        invoke Lambda) across billions of objects; useful for retroactive replication or
        re-encryption after KMS key rotation.
  \item \textbf{Compliance Object Lock}: Compliance mode cannot be overridden by any user
        (including root); satisfies SEC 17a-4, FINRA, HIPAA retention requirements.
\end{itemize}
\end{sapdepth}

\begin{examtip}
\textbf{Common S3 exam traps:}
\begin{itemize}
  \item CRR/SRR requires versioning on \emph{both} buckets --- not just the source.
  \item Replication does NOT copy existing objects --- need S3 Batch Operations.
  \item S3 Gateway Endpoint is \emph{free}; Interface Endpoint (PrivateLink) costs money.
  \item Glacier Deep Archive has 12-hour retrieval --- not ``a few hours'' (use Flexible).
  \item \textbf{S3 is not a file system} --- no append, no atomic file update, no lock.
        For shared file access use EFS; for block storage use EBS.
\end{itemize}
\end{examtip}
